---
title: "220120_tmVac_clean"
output: html_notebook
---

```{r}
print(Sys.time())
```

# Premessa used for pre-processing, and panel editing
quoted out after initial use
```{r}
#suppressMessages(library(premessa))
#paneleditor_GUI()
```

#### Define working directory
```{r setup, include=FALSE}
#This need to be set to where the FCS files are
knitr::opts_knit$set(root.dir = '~/Desktop/')
```

#### Load required packages
```{r}
library(readxl)
library(cowplot)
library(CATALYST)
library(diffcyt)
library(ggplot2) 
library(flowCore)
library(reshape)
library(ggrepel)
library(premessa)
library(flowCore)
library(CytoSpill)
library(plyr)
library(ggpubr)
library(MASS)
library(pheatmap)
library(rstatix)
library(ggrastr)
library(RColorBrewer)
library(viridis)
```

```{r}
sessionInfo()
```

#### Define the seed for the entire workflow
```{r}
my_seed <- 1234
set.seed(my_seed)
```

### import fcs files as a flowset
```{r}
fcs_files <- list.files(pattern = ".fcs$")
fs <- read.flowSet(fcs_files, transformation = FALSE, truncate_max_range = FALSE)
fs
fcs_files
```

### import metadata
```{r}
md <- read_excel("Totalcell_metadata.xlsx")                                  
md
```

### import panel data
```{r}
panel <- "Totalcell_panel.xlsx"                           
panel <- read_excel(panel)
panel 
```

```{r}
# spot check that all panel columns are in the flowSet object
all(panel$fcs_colname %in% colnames(fs))  
```


### Build the SingleCellExperiment from the fcs files, metadata, and panel information
```{r}
sce <- prepData(fs, panel, md)
```

## conduct FlowSOM clustering on the WBC level
```{r}
set.seed(my_seed)
sce <- cluster(sce, 
               features = "type",
               xdim = 10,
               ydim = 10,
               maxK = 30,
               verbose = FALSE, 
               seed = my_seed)   
set.seed(my_seed)
```

## Number of cells measured per sample
```{r, fig.width = 5, fig.height = 2}
plotCounts(sce, 
    group_by = "sample_id", 
    color_by = "condition")
```

## Heatmap of aggregated marker expressions
```{r fig.width = 5}
plotExprHeatmap(sce, 
                features = "type",
                by = "cluster_id", 
                k = "meta20", 
                m = "meta10",
                scale = "first", 
                q = 0.01, 
                perc = T, 
                col_dend = F, 
                bars = T)
```

## Dimensionality reduction (UMAP)
```{r}
set.seed(my_seed)
sce <- runDR(sce, 
             dr = "UMAP", 
             cells = 1000, 
             features = "type",
             n_neighbors = 10) 
set.seed(my_seed)
```

### plot the results coloured by metadata and clustering
```{r, fig.height=4, fig.width=6}
plot_grid(ncol = 2,
  (plotDR(sce, dr = "UMAP", color_by = "patient_id")),
  (plotDR(sce, dr = "UMAP", color_by = "condition")),
  (plotDR(sce, dr = "UMAP", color_by = "sample_id")),
  (plotDR(sce, dr = "UMAP", color_by = "meta20"))
  )
```

## Highlight specific markers expression on the UMAP
```{r, fig.height=6, fig.width=8}
#plotting only type markers
plotDR(sce, 
       dr = "UMAP", 
       color_by = type_markers(sce),
       ncol =6,
       scale = T # set to false if you want to display the unscaled data
       )
```

```{r, fig.height=6, fig.width=8}
#plotting only state markers
plotDR(sce, 
       dr = "UMAP", 
       color_by = state_markers(sce),
       ncol =6,
       scale = T # set to false if you want to display the unscaled data
       )
```

## Merging of clusters into specific cell types
Note! You have to manually make the merging table based on the above clustering
```{r}
merging_table <- read_excel("merging_table1.xlsx")                             
merging_table 
sce <- mergeClusters(sce, k = "meta20", table = merging_table, id = "merging1", overwrite = T)
```

#### Check how your cell type decisions look on the UMAP
```{r, fig.height=4, fig.width=6}
plot_grid(ncol = 2,
  (plotDR(sce, dr = "UMAP", color_by = "meta20")),
  (plotDR(sce, dr = "UMAP", color_by = "merging1"))
  )
```


### plot clusters over the time-course
```{r, fig.width=5, fig.height=2.5}
g = plotAbundances(sce, k = "merging1", by = "sample_id", group_by = "condition")
df = g$data
df$condition = as.factor(df$condition)
df$cluster_id = as.factor(df$cluster_id)

# Multiple line plot
g2 = ggplot(df, aes(x = condition, y = Freq, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, aes(color = cluster_id)) +
  geom_point(aes(fill = cluster_id), size = 1, alpha = 0.6, color = "black", shape = 21) +
  facet_wrap(~cluster_id, ncol = 4, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
g2
```

### UMAP density plot
```{r, fig.height=3,fig.width=4}
p =　plotDR(sce, dr = "UMAP", color_by = "merging1", facet_by = "condition")
p + stat_density_2d(geom = "polygon", contour = TRUE,
      aes(fill = after_stat(level)), colour = "black",
      bins = 8) +
      scale_fill_distiller(palette = "Blues", direction = 1) +
      theme_classic()
```

```{r, fig.height=3,fig.width=4}
#UMAP density plot (nlevel for scaling each plot individually)
p =　plotDR(sce, dr = "UMAP", color_by = "merging1", facet_by = "condition")
pp = p + stat_density_2d(geom = "polygon", contour = TRUE,
      aes(fill = after_stat(nlevel)), colour = "black", size = 0.1, alpha = 0.8,
      bins = 12) +
      scale_fill_distiller(palette = "Spectral", direction = -1) +
      theme_classic()
pp
```


#### Add the info about cell types to the metadata
```{r}
df = as.data.frame(sce@colData$cluster_id)
lookup = sce@metadata$cluster_codes
new <- df  # create a copy of df

# using lapply, loop over columns and match values to the look up table. store in "new".
new[] <- lapply(df, function(x) lookup$merging1[match(x, lookup$som100)])
sce$merging1 = new$`sce@colData$cluster_id`
```

Now we can facet by the new cell type clusters and plot other types of graphs
```{r}
plotDR(sce, dr = "UMAP", color_by = "merging1",facet_by = "merging1")
```

```{r, fig.height=3, fig.width = 3}
plotAbundances(sce, k = "merging1", by = "sample_id", group_by = "condition")
```


### APC re-clustering
```{r}
sce_APC = filterSCE(sce, cluster_id %in% c("mono+DC"), k = "merging1")
rowData(sce_APC)$marker_class = "state"
rowData(sce_APC[c(marker_list1)])$marker_class = "type" 
```

```{r}
set.seed(my_seed)
sce_APC <- cluster(sce_APC, 
               features = "type",
               xdim = 10,
               ydim = 10,
               maxK = 30,
               verbose = FALSE, 
               seed = my_seed)   
set.seed(my_seed)
```

```{r}
set.seed(my_seed)
sce_APC = runDR(sce_APC, 
                  dr = "UMAP", 
                  cells = 1000, 
                  features = type_markers(sce_APC),
                  n_neighbors = 30,
                  repulsion_strength = 1,
                  scale = T
                  ) 
set.seed(my_seed)
```

```{r, fig.height=4}
plot_grid(ncol=3,
plotDR(sce_APC, dr = "UMAP", color_by = "meta10"),
plotDR(sce_APC, dr = "UMAP", color_by = "meta15"),
plotDR(sce_APC, dr = "UMAP", color_by = "meta20")
)
```

```{r, fig.height=4, fig.width=8}
plotDR(sce_APC, 
       dr = "UMAP", 
       color_by = type_markers(sce_APC),
       ncol =5,
       scale = T
       )
```

```{r, fig.height=4, fig.width=8}
plotDR(sce_APC, 
       dr = "UMAP", 
       color_by = state_markers(sce_APC),
       ncol =9,
       scale = T
       )
```

```{r}
merging_table <- read_excel("merging_table_myelo.xlsx")                             
merging_table 
sce_APC <- mergeClusters(sce_APC, k = "meta10", table = merging_table, id = "merging_myelo", overwrite = T)
```

#### Check how your cell type decisions look on the UMAP
```{r, fig.height=4, fig.width=6}
plot_grid(ncol = 2,
  (plotDR(sce_APC, dr = "UMAP", color_by = "meta10")),
  (plotDR(sce_APC, dr = "UMAP", color_by = "merging_myelo"))
  )
```

#### Filtering away DCs
```{r}
sce_mono <- filterSCE(sce_APC, k = "meta10", !cluster_id %in% c(5,6))
```

### monocyte re-clustering
```{r}
rowData(sce_mono)$marker_class = "state"
rowData(sce_mono[c(moi2)])$marker_class = "type"
```

```{r}
set.seed(my_seed)
sce_mono <- cluster(sce_mono, 
               features = "type",
               xdim = 10,
               ydim = 10,
               maxK = 30,
               verbose = FALSE, 
               seed = my_seed)   
set.seed(my_seed)
```

```{r fig.width = 5}
ph = plotExprHeatmap(sce_mono, 
                features = "type",
                by = "cluster_id", 
                k = "meta20", 
                m = "meta10",
                scale = "first", 
                q = 0.01, 
                perc = T, 
                col_dend = F, 
                bars = T)
ph
```

```{r}
set.seed(my_seed)
sce_mono = runDR(sce_mono, 
                  dr = "UMAP", 
                  cells = 1000, 
                  features = type_markers(sce_mono),
                  n_neighbors = 30,
                  scale = T
                  ) 
set.seed(my_seed)
```

## Merging of clusters into specific cell types
Note! You have to manually make the merging table based on the above clustering
```{r}
merging_table <- read_excel("merging_table2.xlsx")                             
merging_table 
sce_mono <- mergeClusters(sce_mono, k = "meta15", table = merging_table, id = "merging2", overwrite = T)
```

#### Check how your cell type decisions look on the UMAP
```{r, fig.height=4, fig.width=6}
plot_grid(ncol = 2,
  (plotDR(sce_mono, dr = "UMAP", color_by = "meta10")),
  (plotDR(sce_mono, dr = "UMAP", color_by = "merging2"))
  )
```

#### Add the info about cell types to the metadata
```{r}
df = as.data.frame(sce_mono@colData$cluster_id)
lookup = sce_mono@metadata$cluster_codes
new <- df  # create a copy of df
new[] <- lapply(df, function(x) lookup$merging2[match(x, lookup$som100)])
sce_mono$merging2 = new$`sce_mono@colData$cluster_id`
```


#### plot the UMAP
```{r}
scCol = c("#E67D72", "#A1A333", "#54BA81")
p = plotDR(sce_mono, dr = "UMAP", color_by = "merging2") +   
  scale_color_manual(values = scCol, name = "subset") +
    theme_minimal(base_size = 14)
p
```

#### rasterize the plot
```{r}
ppp = ggplot(p$data, aes(x, y, colour = merging2, fill = merging2)) +
  geom_point_rast(aes(x, y, colour = merging2, fill = merging2), size =0.3, alpha = 0.8) +
  scale_color_manual(values = scCol, name = "subset") +
  theme_minimal(base_size = 14) + 
  labs(y="UMAP dim1",x = "UMAP dim2") + 
  guides(fill=guide_legend(title="cell types"), colour=guide_legend(title="cell types", override.aes = list(size=5)))
  
  ppp
```

```{r}
myPalette <- colorRampPalette(viridis(11))

p = plotDR(sce_mono, dr = "UMAP", color_by = "CD14") +   
    theme_minimal(base_size = 14)
p
p2 = ggplot(p$data, aes(x, y, colour = value, fill = value)) +
  geom_point_rast(aes(x, y, colour = value, fill = value), size =0.3, alpha = 0.8) +
  theme_minimal(base_size = 14) + 
  scale_colour_gradientn(colours = myPalette(100), limits=c(0, 1)) +
  labs(y="UMAP dim1",x = "UMAP dim2") + 
  guides(fill=guide_legend(title="CD14 scaled"), colour=guide_legend(title="CD14 scaled", override.aes = list(size=5)))
  
p2
```

```{r}
p = plotDR(sce_mono, dr = "UMAP", color_by = "CD16") +   
    theme_minimal(base_size = 14)
p
p3 = ggplot(p$data, aes(x, y, colour = value, fill = value)) +
  geom_point_rast(aes(x, y, colour = value, fill = value), size =0.3, alpha = 0.8) +
  theme_minimal(base_size = 14) + 
  scale_colour_gradientn(colours = myPalette(100), limits=c(0, 1)) +
  labs(y="UMAP dim1",x = "UMAP dim2") + 
  guides(fill=guide_legend(title="CD16 scaled"), colour=guide_legend(title="CD16 scaled", override.aes = list(size=5)))
  
p3
```

```{r, fig.width=6, fig.height=1.5}
plot_grid(ncol=2, p2,p3)
```

### norm to PBMC%
```{r}
sce_PBMC = filterSCE(sce, cluster_id %in% c("B cells", "CD4 T", "CD8 T", "gdT", "Myeloid", "NK cells", "pDC"), k = "merging1")
```

#### extract count data
```{r}
x = sce_mono
shape_by = "sample_id"
group_by = "condition"

ns <- table(
      cluster_id = x$merging2, 
      sample_id = sample_ids(x))

df <- as.data.frame(ns)
    
# add relevant cell metadata
m <- match(df$sample_id, x$sample_id)
for (i in c(shape_by, group_by))
        df[[i]] <- x[[i]][m]
```

#### extract total PBMC count per sample_id
```{r}
df_tot = data.frame(sce_PBMC[["sample_id"]]) # this makes a table for each cell with the available sample ID
y <- table(df_tot) # this sums up the occurrences of each ID. It is thus the actual cell count
df_tot <- data.frame(value = c(y)) #these two lines generates the data frame again to be used for ggplot
df_tot[["sample_id"]] <- rownames(y)
```

#### merge them
```{r}
df_merge = merge(df, df_tot, by = "sample_id")
```

#### add column with percentage of total PBMC
```{r}
df_merge$PBMC_norm = df_merge$Freq/df_merge$value*100
```

#### plot PBMC norm percentage
```{r, fig.width=5, fig.height=1.5}
g2 = ggplot(df_merge, aes(x = condition, y = PBMC_norm, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, aes(color = cluster_id)) +
  geom_point(aes(fill = cluster_id), size = 1, alpha = 0.6, color = "black", shape = 21) +
    facet_wrap(~cluster_id, ncol = 3, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
g2 + expand_limits(y=0)
```
### stats
```{r}
#first subset to one cell type
df_merge_sub = subset(df_merge, cluster_id == "cMono")

# paired t test
stat.test <- pairwise_t_test(PBMC_norm ~ condition, data = df_merge_sub, paired = TRUE, p.adjust.method = "none") 
stat.test
```

```{r}
#narrow down to the test that should be displayed
stat.test_sub = stat.test[c(1,16),]
stat.test_sub
```

```{r, fig.width=5, fig.height=1.5}
g1 = ggplot(df_merge_sub, aes(x = condition, y = PBMC_norm, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, color = scCol[1]) +
  geom_point(fill = scCol[1], size = 1, alpha = 0.6, color = "black", shape =21) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + 
  expand_limits(y=c(0, max(df_merge_sub$PBMC_norm)+5)) +
  stat_pvalue_manual(stat.test_sub, label = "p.adj.signif", y.position = c(max(df_merge_sub$PBMC_norm)+3, max(df_merge_sub$PBMC_norm)+3)) +
  ggtitle("cMono") +
  ylab("% of PBMCs") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g1
```

```{r}
#first subset to one cell type
df_merge_sub = subset(df_merge, cluster_id == "ncMono")

# paired t test
stat.test <- pairwise_t_test(PBMC_norm ~ condition, data = df_merge_sub, paired = TRUE, p.adjust.method = "none") 
stat.test
```

```{r}
#narrow down to the test that should be displayed
stat.test_sub = stat.test[c(1,16),]
stat.test_sub
```

```{r, fig.width=5, fig.height=1.5}
g2 = ggplot(df_merge_sub, aes(x = condition, y = PBMC_norm, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, color = scCol[3]) +
  geom_point(fill = scCol[3], size = 1, alpha = 0.6, color = "black", shape =21) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + 
  expand_limits(y=c(0, max(df_merge_sub$PBMC_norm)+1)) +
  stat_pvalue_manual(stat.test_sub, label = "p.adj.signif", y.position = c(max(df_merge_sub$PBMC_norm)+0.5, max(df_merge_sub$PBMC_norm)+0.5)) +
  ggtitle("ncMono") +
  ylab("% of PBMCs") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g2
```

```{r}
#first subset to one cell type
df_merge_sub = subset(df_merge, cluster_id == "intMono")

# paired t test
stat.test <- pairwise_t_test(PBMC_norm ~ condition, data = df_merge_sub, paired = TRUE, p.adjust.method = "none") 
stat.test
```

```{r}
#narrow down to the test that should be displayed
stat.test_sub = stat.test[c(1,16),]
stat.test_sub
```

```{r, fig.width=5, fig.height=1.5}
g3 = ggplot(df_merge_sub, aes(x = condition, y = PBMC_norm, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, color = scCol[2]) +
  geom_point(fill = scCol[2], size = 1, alpha = 0.6, color = "black", shape =21) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + 
  expand_limits(y=c(0, max(df_merge_sub$PBMC_norm)+0.3)) +
  stat_pvalue_manual(stat.test_sub, label = "p.adj.signif", y.position = c(max(df_merge_sub$PBMC_norm)+0.2, max(df_merge_sub$PBMC_norm)+0.2)) +
  ggtitle("intMono") +
  ylab("% of PBMCs") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g3
```




```{r}
plot_grid(ncol=3, g1, g2, g3)
```

### Feature of interest over time
```{r}
test1  = as.data.frame(t(sce_mono@assays@data$exprs))
test1$sample_id = sce_mono$sample_id
test1$condition = sce_mono$condition
test1$patient_id = sce_mono$patient_id
test1$cluster_id = sce_mono$merging2
test1$combo_id = paste0(test1$sample_id, "_", test1$cluster_id)
test2 = ddply(test1, .(combo_id), summarize,  HLADR_mean=mean(get("HLA-DR")))
```

```{r}
marker_list = rownames(sce)
marker_list_mean = c()
  for (i in 1:length(marker_list)) {
  marker_list_mean[i] = paste(marker_list[i], "mean",sep="_")
  }
```

```{r}
test4 = data.frame(combo_id = test2$combo_id) 

for(i in c(1:length(marker_list))) {
  nam = paste0("ttest",i) 
  nam2 = marker_list[i]
    g = ddply(test1, .(combo_id), summarize, mean((get(nam2))))
  assign(nam, g)
  new <- rep(i, nrow(test4))
  test4[ , ncol(test4) + 1] <- g[2] 
}
```

```{r}
colnames(test4) = c("combo_id", marker_list_mean)
```

```{r}
test3 = merge(test4, test1, by = "combo_id")
test3 = test3[!duplicated(test3$combo_id),]
test3
```

```{r}
test3$combo_id2 = paste0(test3$condition, "_", test3$cluster_id)
```


```{r}
for(i in c(1:length(marker_list))) {
df_sub = test3[,c(1,i+1,88:92)]
nam = paste0("df_list_",i) 
assign(nam, melt(df_sub))
}
```

```{r}
#making a list of lists.
df_list_list = list()
for(i in c(1:length(marker_list))) {
  df_list_list[i] = list(get(paste0("df_list_",i)))
}
```

```{r, fig.width=5, fig.height=1.5}
for(i in c(1:length(marker_list))) {
  nam = paste0("g",i) 
  ggg = ggplot(data = df_list_list[[i]],
    aes(x = condition, y = value, group = cluster_id)) +
    stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
    stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, aes(color = cluster_id)) +
    geom_point(aes(fill = cluster_id), size = 1, alpha = 0.6, color = "black", shape = 21) +
    facet_wrap(~cluster_id, ncol = 3, scales = "free") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+ 
    expand_limits(y=c(0, 0.5))  +
    labs(y=paste(marker_list_mean[i], "expression"), x = "")
  
  assign(nam, ggg)
}
```

```{r}
#making a list of lists.
g_list = list()
for(i in c(1:length(marker_list))) {
  g_list[i] = list(get(paste0("g",i)))
}
```


```{r, fig.width=24, fig.height=8}
gridExtra::grid.arrange(grobs = g_list)
```



## make a graph with stats instead of curves
It is not possible to run stats_summary + stat_compare_means together, as they depend on different groups. So can only make a plot of the median line (stats summary) or one with the statistical significance (stat compare means). facetting is ok!

```{r, fig.width=5, fig.height=1.5}
my_comparisons <- list(c("day00", "day01"), c("day20", "day22"))

for(i in c(1:length(marker_list))) {
  nam = paste0("g",i) 
  ggg = ggplot(data = df_list_list[[i]],
    aes(x = condition, y = value, group = condition)) +
    geom_point(aes(fill = cluster_id), size = 1, alpha = 0.6, color = "black", shape = 21) +
    facet_wrap(~cluster_id, ncol = 3, scales = "free") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+ 
    expand_limits(y=c(0, 0.5))  +
    labs(y=paste(marker_list_mean[i], "expression"), x = "")
  
    gggg = ggg + stat_compare_means(aes(group = condition), comparisons = my_comparisons, method = "t.test", paired = T, label = "p.signif", vjust = 0.6)

  assign(nam, gggg)
}
```

```{r}
#making a list of lists.
g_list = list()
for(i in c(1:length(marker_list))) {
  g_list[i] = list(get(paste0("g",i)))
}
```

```{r, fig.width=24, fig.height=8}
gridExtra::grid.arrange(grobs = g_list)
```


### Make select marker graphs with cell types next to eachother and paired t test
```{r}
test4 = data.frame(combo_id = test2$combo_id) 

for(i in c(1:length(marker_list))) {
  nam = paste0("ttest",i) 
  nam2 = marker_list[i]
    g = ddply(test1, .(combo_id), summarize, mean((get(nam2))))
  assign(nam, g)
  new <- rep(i, nrow(test4))
  test4[ , ncol(test4) + 1] <- g[2] 
}
```

#### rename marker list items 
```{r}
marker_list_mean
marker_list_mean[2] = "markoi1" #marker of interest
marker_list_mean[5] = "markoi2"
marker_list_mean[6] = "markoi3"
```


```{r}
colnames(test4) = c("combo_id", marker_list_mean)
```


```{r}
test3 = merge(test4, test1, by = "combo_id")

test3 = test3[!duplicated(test3$combo_id),]
test3
```




```{r}
#first subset to one cell type
df_merge_sub = subset(test3, cluster_id == "cMono")
#df_merge_sub = df_merge_sub[,c(3,88:91)]

# paired t test
stat.test <- pairwise_t_test(markoi1 ~ condition, data = df_merge_sub, paired = TRUE, p.adjust.method = "none") 
stat.test
```

```{r}
#narrow down to the test that should be displayed
stat.test_sub = stat.test[c(1,16),]
stat.test_sub
```

```{r, fig.width=5, fig.height=1.5}
g1 = ggplot(df_merge_sub, aes(x = condition, y = markoi1, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, color = scCol[1]) +
  geom_point(fill = scCol[1], size = 1, alpha = 0.6, color = "black", shape =21) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + 
  expand_limits(y=c(0, max(df_merge_sub$markoi1)+0.1)) +
  stat_pvalue_manual(stat.test_sub, label = "p.adj.signif", y.position = c(max(df_merge_sub$markoi1)+0.05, max(df_merge_sub$markoi1)+0.05)) +
  ggtitle("cMono") +
  #ylab("% of PBMCs") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g1
```

```{r}
#first subset to one cell type
df_merge_sub = subset(test3, cluster_id == "ncMono")
#df_merge_sub = df_merge_sub[,c(3,88:91)]

# paired t test
stat.test <- pairwise_t_test(markoi1 ~ condition, data = df_merge_sub, paired = TRUE, p.adjust.method = "none") 
stat.test
```

```{r}
#narrow down to the test that should be displayed
stat.test_sub = stat.test[c(1,16),]
stat.test_sub
```

```{r, fig.width=5, fig.height=1.5}
g2 = ggplot(df_merge_sub, aes(x = condition, y = markoi1, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, color = scCol[3]) +
  geom_point(fill = scCol[3], size = 1, alpha = 0.6, color = "black", shape =21) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + 
  expand_limits(y=c(0, max(df_merge_sub$markoi1)+0.2)) +
  stat_pvalue_manual(stat.test_sub, label = "p.adj.signif", y.position = c(max(df_merge_sub$markoi1)+0.1, max(df_merge_sub$markoi1)+0.1)) +
  ggtitle("ncMono") +
  #ylab("% of PBMCs") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g2
```


```{r}
#first subset to one cell type
df_merge_sub = subset(test3, cluster_id == "intMono")
#df_merge_sub = df_merge_sub[,c(3,88:91)]

# paired t test
stat.test <- pairwise_t_test(markoi1 ~ condition, data = df_merge_sub, paired = TRUE, p.adjust.method = "none") 
stat.test
```

```{r}
#narrow down to the test that should be displayed
stat.test_sub = stat.test[c(1,16),]
stat.test_sub
```

```{r, fig.width=5, fig.height=1.5}
g3 = ggplot(df_merge_sub, aes(x = condition, y = markoi1, group = cluster_id)) +
  stat_summary(fun=median, geom="line", size = 1.5, alpha = 0.7, color = "black") +
  stat_summary(fun=median, geom="line", size = 1.1, alpha = 1, color = scCol[2]) +
  geom_point(fill = scCol[2], size = 1, alpha = 0.6, color = "black", shape =21) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + 
  expand_limits(y=c(0, max(df_merge_sub$markoi1)+0.2)) +
  stat_pvalue_manual(stat.test_sub, label = "p.adj.signif", y.position = c(max(df_merge_sub$markoi1)+0.1, max(df_merge_sub$markoi1)+0.1)) +
  ggtitle("intMono") +
  #ylab("% of PBMCs") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
g3
```

```{r}
plot_grid(ncol=3, g1, g2, g3)
```

Repeat for other markers of interest


## “Featureplot” over time for selected markers
```{r, fig.height=4, fig.width=8}
up1 = plot_grid(ncol = 1,
plotDR(sce_mono, dr = "UMAP", color_by = "markoi1", ncol =7, scale = T, facet_by = "condition"),
plotDR(sce_mono, dr = "UMAP", color_by = "markoi2", ncol =7, scale = T, facet_by = "condition"),
plotDR(sce_mono, dr = "UMAP", color_by = "markoi3", ncol =7, scale = T, facet_by = "condition")
)
up1
```

```{r, fig.height=4, fig.width=8}
up2 = plot_grid(ncol = 1,
plotDR(sce_mono, dr = "UMAP", color_by = "markoi1", ncol =7, scale = F, facet_by = "condition"),
plotDR(sce_mono, dr = "UMAP", color_by = "markoi2", ncol =7, scale = F, facet_by = "condition"),
plotDR(sce_mono, dr = "UMAP", color_by = "markoi3", ncol =7, scale = F, facet_by = "condition")
)
up2
```

```{r}
print(Sys.time())
```
